\bf Formatting Your Paper \\ for the \confname~Conference

### 2016-06-10 acl.2016/amta2016.tex
cite format

For many users, decoding is the most time-critical part of the translation process. Making use of the multiple cores that are now ubiquitous in todays servers is a common strategy to ameliorate this issue. However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\cite{mfernandez2016boosting}. That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores. There have been speculation on the causes of the inefficiency as well as potential remedies.

@@

~\citep{Heafield-kenlm} and ~\citep{yasuhara-EtAl:2013:EMNLP} describes fast, efficient datastructures for language models. ~\citep{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements. ~\citep{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.

~\citep{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm which allows the tradeoff between speed and translation quality to the adjusted with a single parameter. ~\citep{wuebker2012fast} note that language model querying is amongst the most expensive operation in decoding. They sought to improved decoding speed by caching score computations early pruning of translation options. This work is similar to ~\citep{Heafield-mtplz} which group hypotheses with identical language model context and incrementally expand them, reducing LM querying.

~\citep{mfernandez2016boosting} was concerned with multi-core speed but treated the decoding process as a black box within a parallelization framework.

@@

The most well known is Moses~\cite{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning. It is widely used for MT research and commercial use.

Joshua~\cite{Joshua-Decoder} also supports cube-pruning for phrase-based models. Phrasal~\cite{spence2014phrasal} supports a number of variants of the phrase-based model. Jane~\cite{peitz2012jane} supports the language model look-ahead described in ~\citep{wuebker2012fast} in addition to other tools to speed up decoding such as having a separate fast, lightweight decoder. mtplz is a specialized decoder developed to implement the incremental decoding described in ~\citep{Heafield-mtplz}.

@@

%The first four feature functions are simple and do not leave much room for optimization. As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\citep{Heafield-kenlm} and ~\citep{yasuhara-EtAl:2013:EMNLP}.

@@

Libraries such as tcmalloc~\cite{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.

@@

However, there are alternatives to this configuration. ~\citep{Och:2001a} uses a single stack for all hypotheses, ~\citep{Brown:1993} uses coverage stacks (ie. one stack per unique coverage vector) while ~\citep{peitz2012jane} and ~\citep{Zens+Ney:2008:iwslt} apply both coverage and cardinality pruning. While useful, these prior works present only one particular stack configuration each.

~\citep{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.

@@

For any phrase-table table of a realistic size, memory and loading time constraints requires us to use a load-on-demand implementation. Moses has several which we can make use of, each with differing performance characteristics. Figure~\ref{fig:moses-phrase-tables-time} shows the decoding speed for the fastest two implementations.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
\end{tabular}
\caption{Moses decoding speed with two different phrase-table implementations}
\label{fig:moses-phrase-tables-time}
\end{figure} 
From this, it appears that the Probing phrase-table~\cite{Bogoychev.probing.2016} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.

@@

Secondly, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule. Compression was championed by ~\citep{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads. We shall therefore take the opposite approach to and improve decoding speed by disabling compression.

@@

Similar to the phrase-table, the lexicalized re-ordering model is trained on parallel data. A resultant model file is then queried during decoding. The need for random lookup during querying inevitably results in slower decoding speed. Previous work such as~\citep{junczys_tsd_2012b} improve querying speed with more efficient data structures.

@@

This optimization has precedent in ~\citep{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.

@@

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure. These binary formats were chosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}

@@

Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\cite{Chiang:2007:cl} with a pop-limit of 400\footnote{the pop-limit was chosen from public discussion on the Moses mailing list on an acceptable balance between decoding speed and translation quality with Moses for commercial use %https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html 
}, rather than the basic phrase-based algorithm. The cube-pruning algorithm is often employed by users who require fast decoding as it gives them the ability to trade speed with translation quality via a simple pop-limit parameter.

@@

%Table~\ref{tab:stack-configuration} and 
Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits. The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits. It is also slower but the time/quality tradeoff is better overall with this stack configuration. Model scores for coverage and cardinality configurations are identical, which differs from the results in ~\citep{ortizmartinez-garciavarea-casacuberta:2006:WMT}. This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.
%\begin{table*}[t]
%\small
%\begin{center}
%\begin{tabular}{|r|r|r|r|r|r|r|} \hline
%Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
%		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
%100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
%500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
%1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
%5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
%10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
%15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
%20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
%\end{tabular}
%\end{center}
%\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
%\label{tab:stack-configuration}
%\end{table*}

@@

%We mimic the use of multi-core servers by single-threaded decoders such as mtplz~\cite{Heafield-mtplz} and Jane~\cite{peitz2012jane} by running multiple processes of the Moses decoder, each running one thread. The input is distributed to the processes using a external script. The result in Figure~\ref{fig:moses-speed} (dashed line) shows that this method scales much better than multithreading the application.

>>>>

For many users, decoding is the most time-critical part of the translation process. Making use of the multiple cores that are now ubiquitous in todays servers is a common strategy to ameliorate this issue. However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\citep{mfernandez2016boosting}. That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores. There have been speculation on the causes of the inefficiency as well as potential remedies.

@@

~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP} describes fast, efficient datastructures for language models. ~\cite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements. ~\cite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.

~\cite{Chiang:2007:cl} describes the cube-pruning and cube-growing algorithm which allows the tradeoff between speed and translation quality to the adjusted with a single parameter. ~\cite{wuebker2012fast} note that language model querying is amongst the most expensive operation in decoding. They sought to improved decoding speed by caching score computations early pruning of translation options. This work is similar to ~\cite{Heafield-mtplz} which group hypotheses with identical language model context and incrementally expand them, reducing LM querying.

~\cite{mfernandez2016boosting} was concerned with multi-core speed but treated the decoding process as a black box within a parallelization framework.

@@

The most well known is Moses~\citep{koehn-EtAl:2007:PosterDemo} which implements a number of speed optimizations, including cube-pruning. It is widely used for MT research and commercial use.

Joshua~\citep{Joshua-Decoder} also supports cube-pruning for phrase-based models. Phrasal~\citep{spence2014phrasal} supports a number of variants of the phrase-based model. Jane~\citep{peitz2012jane} supports the language model look-ahead described in ~\cite{wuebker2012fast} in addition to other tools to speed up decoding such as having a separate fast, lightweight decoder. mtplz is a specialized decoder developed to implement the incremental decoding described in ~\cite{Heafield-mtplz}.

@@

%The first four feature functions are simple and do not leave much room for optimization. As have already been mentioned, the language model consumes a large part of decoding time but it has already been heavily optimized in previous works such as ~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP}.

@@

Libraries such as tcmalloc~\citep{ghemawat2009tcmalloc} are designed to reduce locking contention for multi-threaded application but in our case, this is still not enough.

@@

However, there are alternatives to this configuration. ~\cite{Och:2001a} uses a single stack for all hypotheses, ~\cite{Brown:1993} uses coverage stacks (ie. one stack per unique coverage vector) while ~\cite{peitz2012jane} and ~\cite{Zens+Ney:2008:iwslt} apply both coverage and cardinality pruning. While useful, these prior works present only one particular stack configuration each.

~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.

@@

For any phrase-table table of a realistic size, memory and loading time constraints requires us to use a load-on-demand implementation. Moses has several which we can make use of, each with differing performance characteristics. Figure~\ref{fig:moses-phrase-tables-time} shows the decoding speed for the fastest two implementations.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
\end{tabular}
\caption{Moses decoding speed with two different phrase-table implementations}
\label{fig:moses-phrase-tables-time}
\end{figure} 
From this, it appears that the Probing phrase-table~\citep{Bogoychev.probing.2016} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.

@@

Secondly, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule. Compression was championed by ~\cite{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads. We shall therefore take the opposite approach to and improve decoding speed by disabling compression.

@@

Similar to the phrase-table, the lexicalized re-ordering model is trained on parallel data. A resultant model file is then queried during decoding. The need for random lookup during querying inevitably results in slower decoding speed. Previous work such as~\cite{junczys_tsd_2012b} improve querying speed with more efficient data structures.

@@

This optimization has precedent in ~\cite{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.

@@

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\citep{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All model files were then binarized; the language models were binarized using KenLM~\citep{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure. These binary formats were chosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}

@@

Standard Moses phrase-based configurations are used, except that we use the cube-pruning algorithm~\citep{Chiang:2007:cl} with a pop-limit of 400\footnote{the pop-limit was chosen from public discussion on the Moses mailing list on an acceptable balance between decoding speed and translation quality with Moses for commercial use %https://www.mail-archive.com/moses-support\%40mit.edu/msg13181.html 
}, rather than the basic phrase-based algorithm. The cube-pruning algorithm is often employed by users who require fast decoding as it gives them the ability to trade speed with translation quality via a simple pop-limit parameter.

@@

%Table~\ref{tab:stack-configuration} and 
Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model at various pop-limits. The model scores for all stack configurations are identical for low pop-limits but grouping hypotheses into coverage \& end position produces higher model scores for higher pop-limits. It is also slower but the time/quality tradeoff is better overall with this stack configuration. Model scores for coverage and cardinality configurations are identical, which differs from the results in ~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}. This may be due to differing test set characteristics, or interactions with cube-pruning algorithm we used.
%\begin{table*}[t]
%\small
%\begin{center}
%\begin{tabular}{|r|r|r|r|r|r|r|} \hline
%Pop-limit	& \multicolumn{2}{c|}{Cardinality}	& \multicolumn{2}{c|}{Coverage}	& \multicolumn{2}{c|}{Coverage \& end pos} \\ \hline
%		& Time		& Score			& Time		& Score		& Time		& Score	\\ \hline
%100  		& 73		& -8.64513 		& 75		& -8.64513	& 72	& -8.64513 \\
%500		& 237		& -8.59563		& 225		& -8.59563 	& 229	& -8.59612 \\
%1,000	 	& 416		& -8.58700		& 397		& -8.58700	& 423	& -8.58700 \\ 
%5,000 		& 1930		& -8.58165		& 1931		& -8.58165	& 2153	& -8.58098 \\ 
%10,000 		& 3619		& -8.58133		& 3630		& -8.58133	& 4576	& -8.58015 \\ 
%15,000 		& 4830		& -8.58130		& 5001		& -8.58130	& 7156	& -8.57999 \\ 
%20,000 		& 5849		& -8.58130		& 5916		& -8.58130	& 9583	& -8.57994 \\ \hline
%\end{tabular}
%\end{center}
%\caption{Decoding time (in secs with 32 threads) and average model scores for different stack configurations}
%\label{tab:stack-configuration}
%\end{table*}

@@

%We mimic the use of multi-core servers by single-threaded decoders such as mtplz~\citep{Heafield-mtplz} and Jane~\citep{peitz2012jane} by running multiple processes of the Moses decoder, each running one thread. The input is distributed to the processes using a external script. The result in Figure~\ref{fig:moses-speed} (dashed line) shows that this method scales much better than multithreading the application.
